---
title: "Compiling Data Sources"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(terra)
library(dbplyr)
library(DBI)
library(tidyverse)
library(bcinvadeR)
library(bcdata)

base_dir = stringr::str_extract(getwd(),"C:\\/Users\\/[a-zA-Z]+")
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")
lan_root = "//SFP.IDIR.BCGOV/S140/S40203/RSD_ FISH & AQUATIC HABITAT BRANCH/General/"

requery_occs = F

# Custom function to show number of rows with commas.
nrowc = function(dat){
  scales::comma(nrow(dat))
}
```

### Data Sources: {.tabset}

  - HydroClim (https://www.earthenv.org/streams). Authors used a 30 arc sec hydrograph that describes streams, rivers, and perhaps lakes, to trim WorldClim **air** temperature to the geometries of these waterbodies (full paper explaining this data source: https://www.nature.com/articles/sdata201573#Sec2)
  - Environmental Monitoring Database (EMS) Water Temperature Records - hosted on BC Data Catalogue.
  - Occurrence data for fish species that are sensitive to Whirling Disease: Bull Trout, Sockeye Salmon, others?
  - Fish Life-stage History / Seasonality

#### Hydroclim
```{r get_hydroclim}
hydroclim_layers_in_onedrive = list.files(
  path = paste0(onedrive_wd,"raster/hydroclim_goldMussel_rast/"),
  pattern = "^bc_.*\\.tif$",
  full.names = T
)

hydroclim_layer_names = paste0(str_remove(str_replace_all(str_extract(hydroclim_layers_in_onedrive,"bc_.*\\.tif$"),"(_|\\.)"," "),"( rast)? tif"), collapse = ', ')

hydroclim = terra::rast(hydroclim_layers_in_onedrive[1])

terra::plot(hydroclim)
```

Layers currently available based on Hydroclim data: `r hydroclim_layer_names`.

#### EMS Water Temperature
```{r}

# The following path will work on Chris' computer. Not on Johns, though.
path_to_EMS = paste0(base_dir,"/Downloads/LocalR/long_term_projects/EMS/output/EMS.sqlite")

if(file.exists(path_to_EMS)){
  ems_db = DBI::dbConnect(RSQLite::SQLite(), path_to_EMS)
}

DBI::dbListTables(ems_db)

# Query the EMS database for water temperature records
ems_wt = DBI::dbGetQuery(ems_db, "select * from results where PARAMETER like 'Temperature' or PARAMETER like 'Temperature-Field'")

# Convert the date field from character to datetime.
ems_wt = ems_wt |> 
  dplyr::mutate(the_date = lubridate::ymd_hms(COLLECTION_START)) |> 
  dplyr::mutate(the_year = lubridate::year(the_date))
```

The EMS database has `r nrowc(ems_wt)` rows, where `r nrowc(ems_wt[ems_wt$PARAMETER == 'Temperature',])` rows are of parameter type 'Temperature' and `r nrowc(ems_wt[ems_wt$PARAMETER == 'Temperature-Field',])` rows are of parameter type 'Temperature-Field.'

Number of rows of EMS temperature data by parameter and year:
```{r ems_table}
### May need to filter by the source
ems_wt |> 
  sf::st_drop_geometry() |> 
  dplyr::count(PARAMETER,the_year) |> 
  tidyr::pivot_wider(names_from = PARAMETER, values_from = n, values_fill = 0) |> 
  DT::datatable()
```

Plots of number of EMS data points by grid cell:
```{r ems_example_pictures}
bc = bcmaps::bc_bound() |> 
  sf::st_transform(4326)

bc_g = sf::st_make_grid(bc, n = c(50,50)) |> 
  sf::st_as_sf() |> 
  dplyr::mutate(cell_id = dplyr::row_number())

ems_sf = ems_wt |> 
  dplyr::filter(!is.na(LONGITUDE)) |> 
  sf::st_as_sf(coords = c("LONGITUDE","LATITUDE"), crs = 4326)

ems_sf = ems_sf |> 
  sf::st_join(bc_g)

ems_data_by_cell = ems_sf |> 
  sf::st_drop_geometry() |>
  dplyr::count(PARAMETER,cell_id)

bc_g_w_dat = bc_g |> 
  dplyr::left_join(ems_data_by_cell) |> 
  dplyr::filter(!is.na(n))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = bc_g, fill = 'transparent', col = 'grey') + 
  geom_sf(data = bc_g_w_dat, aes(fill = n)) + 
  facet_wrap( ~ PARAMETER)
```

#### Fish Occurrence Records
```{r}
# Fish occurrence records queried from typical data sources.
all_occs = sf::read_sf(paste0(lan_root,"2 SCIENCE - Invasives/SPECIES/5_Incidental Observations/AIS_occurrences_all_sources.gpkg"))

# rel_occs = all_occs |> 
#   dplyr::filter(Species %in% c("Bull trout","Sockeye salmon"))

# Add queries for Bull Trout and Sockeye Salmon
if(requery_occs){
  bt_occs = suppressMessages(suppressWarnings(grab_aq_occ_data('Bull Trout', quiet = T)))
  ss_occs = suppressMessages(suppressWarnings(grab_aq_occ_data('Sockeye Salmon')))
  
  sf::write_sf(bt_occs, "data/bull_trout_occurrences.gpkg")
  sf::write_sf(ss_occs, "data/sockeye_salmon_occurrences.gpkg")
} else {
  bt_occs = sf::read_sf("data/bull_trout_occurrences.gpkg")
  ss_occs = sf::read_sf("data/sockeye_salmon_occurrences.gpkg")
}

fish_occs = dplyr::bind_rows(bt_occs, ss_occs)

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = fish_occs) +
  facet_wrap( ~ Species)
```

#### Fish Periodicity

I'm just waiting on Amy Sigvaldson's work to amalgamate edits from Ron Ptolemy, DFO, and other groups to finalize the fish periodicity file. This file describes when species' life stages occur in major rivers.
