---
title: "Compiling Data Sources"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(terra)
library(dbplyr)
library(DBI)
library(tidyverse)
library(bcinvadeR)
library(bcdata)
library(leaflet)
library(leafpop)
library(htmltools)

base_dir = stringr::str_extract(getwd(),"C:\\/Users\\/[a-zA-Z]+")
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")
lan_root = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/"

requery_occs = F

# Custom function to show number of rows with commas.
nrowc = function(dat){
  scales::comma(nrow(dat))
}
```

### Data Sources: {.tabset}

  - HydroClim (https://www.earthenv.org/streams). Authors used a 30 arc sec hydrograph that describes streams, rivers, and perhaps lakes, to trim WorldClim **air** temperature to the geometries of these waterbodies (full paper explaining this data source: https://www.nature.com/articles/sdata201573#Sec2)
  - Environmental Monitoring Database (EMS) Water Temperature Records - hosted on BC Data Catalogue.
  - Occurrence data for fish species that are sensitive to Whirling Disease: Bull Trout, Sockeye Salmon, others?
  - Fish Life-stage History / Seasonality

#### Watershed Groups

These are the spatial units that we use to filter the province for abiotic conditions. There are 246 of them in BC.

```{r}
if(!file.exists("data/freshwater_atlas_named_watersheds.gpkg")){
  ws = bcdata::bcdc_query_geodata('freshwater-atlas-watershed-groups') |> # 11,580
    bcdata::collect()
  ws = sf::st_transform(ws, 4326)
  sf::write_sf(ws, "data/freshwater_atlas_watershed_groups.gpkg")
  ws = sf::st_simplify(ws)
  sf::write_sf(ws, "data/freshwater_atlas_watershed_groups_simplified.gpkg")
} else {
  ws = sf::read_sf("data/freshwater_atlas_watershed_groups_simplified.gpkg")
}

ggplot() + 
  geom_sf(data = ws)
```

#### Hydroclim
```{r get_hydroclim}
hydroclim_layers_in_onedrive = list.files(
  path = paste0(onedrive_wd,"raster/hydroclim_goldMussel_rast/"),
  pattern = "^bc_.*\\.tif$",
  full.names = T
)

hydroclim_layer_names = paste0(str_remove(str_replace_all(str_extract(hydroclim_layers_in_onedrive,"bc_.*\\.tif$"),"(_|\\.)"," "),"( rast)? tif"), collapse = ', ')

hydroclim = terra::rast(hydroclim_layers_in_onedrive[1])

terra::plot(hydroclim)
```

Layers currently available based on Hydroclim data: `r hydroclim_layer_names`.

#### EMS Water Temperature
```{r ems_water}

# The following path will work on Chris' computer. Not on Johns, though.
if(str_detect(onedrive_wd,"CMADSEN")){
  path_to_EMS = paste0(base_dir,"/Downloads/LocalR/long_term_projects/EMS/output/EMS.sqlite")
} else {
  path_to_EMS <- file.path(dirname(onedrive_wd), "R_projects", "EMS", "output", "EMS.sqlite")
}

if(file.exists(path_to_EMS)){
  ems_db = DBI::dbConnect(RSQLite::SQLite(), path_to_EMS)
}

DBI::dbListTables(ems_db)

# Query the EMS database for water temperature records
ems_wt = DBI::dbGetQuery(ems_db, "select * from results where PARAMETER like 'Temperature' or PARAMETER like 'Temperature-Field'")

# Convert the date field from character to datetime.
ems_wt = ems_wt |> 
  dplyr::mutate(the_date = lubridate::ymd_hms(COLLECTION_START)) |> 
  dplyr::mutate(the_year = lubridate::year(the_date))


ems_wt<-ems_wt |> 
    dplyr::filter(!is.na(RESULT)) %>% 
    dplyr::filter(LOCATION_TYPE == "RIVER,STREAM OR CREEK" |
                    LOCATION_TYPE == "MONITORING WELL" |
                    LOCATION_TYPE == "LAKE OR POND")
```

The EMS database, was filtered for locations that were tagged as "River, Stream, Or Creek", "Monitoring Well", or " Lake or Pond". The data has `r nrowc(ems_wt)` rows, where `r nrowc(ems_wt[ems_wt$PARAMETER == 'Temperature',])` rows are of parameter type 'Temperature' and `r nrowc(ems_wt[ems_wt$PARAMETER == 'Temperature-Field',])` rows are of parameter type 'Temperature-Field.'

Number of rows of EMS temperature data by parameter and year:
```{r ems_table}
### May need to filter by the source
ems_wt |> 
  sf::st_drop_geometry() |> 
  dplyr::count(PARAMETER,the_year) |> 
  tidyr::pivot_wider(names_from = PARAMETER, values_from = n, values_fill = 0) |> 
  DT::datatable()



### Temperatures where there are a lot of readings look good - no big values outside what could be considered a normal range
# ems_wt |> 
#   sf::st_drop_geometry() |> 
#   dplyr::count(PARAMETER, MONITORING_LOCATION) |> 
#   tidyr::pivot_wider(names_from = PARAMETER, values_from = n, values_fill = 0) |> 
#   DT::datatable()
# 
# 
# top_locations <- ems_wt |>
#   sf::st_drop_geometry() |> 
#   count(MONITORING_LOCATION, sort = TRUE) |> 
#   slice_head(n = 10) |> 
#   pull(MONITORING_LOCATION)
#   
# ems_wt_top10 <- ems_wt |> 
#   filter(MONITORING_LOCATION %in% top_locations)
# 
# ggplot(ems_wt_top10, aes(x = COLLECTION_DATE, y = RESULT, color = PARAMETER)) +
#   geom_line() +
#   geom_point(alpha = 0.5) +
#   facet_wrap(~ MONITORING_LOCATION, scales = "free_y") +
#   labs(
#     title = "Temperature Readings Over Time at Top 10 Monitoring Locations",
#     x = "Date",
#     y = "Temperature (units)",  # Replace with actual units if available
#     color = "Parameter"
#   ) +
#   theme_minimal()



```

Plots of number of EMS data points by grid cell:
```{r ems_example_pictures}
bc = bcmaps::bc_bound() |> 
  sf::st_transform(4326)

bc_g = sf::st_make_grid(bc, n = c(50,50)) |> 
  sf::st_as_sf() |> 
  dplyr::mutate(cell_id = dplyr::row_number())

ems_sf = ems_wt |> 
  dplyr::filter(!is.na(LONGITUDE)) |> 
  sf::st_as_sf(coords = c("LONGITUDE","LATITUDE"), crs = 4326)

ems_sf = ems_sf |> 
  sf::st_join(bc_g)

ems_data_by_cell = ems_sf |> 
  sf::st_drop_geometry() |>
  dplyr::count(PARAMETER,cell_id)

bc_g_w_dat = bc_g |> 
  dplyr::left_join(ems_data_by_cell) |> 
  dplyr::filter(!is.na(n))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = bc_g, fill = 'transparent', col = 'grey') + 
  geom_sf(data = bc_g_w_dat, aes(fill = n)) + 
  facet_wrap( ~ PARAMETER)
```


#### BC Aquarius Data

The BC Aquarius data is comprised of a range of temperature readings from across the province. There are a lot of data, with varying temporal scales. Information on the purpose of the readings is limited. This could be a great resource for waterbody specific data, where we can see trends in the waterbody's temperature and link this to key points in fish, tubifex and M. cerebralis life stages. 

I am waiting on API access, then we should be able to query the data. Currently we can download data from: https://www.env.gov.bc.ca/wsd/data_searches/water/

Data has just been averaged, with min and max temperatures also recorded. The dates for each of the locations are available in the leaflet map, when hovering.

```{r get_aquarius_data}

aqua_data<-read.csv("data/WaterTemperature.csv")

names(aqua_data) = gsub("\\.", "_", names(aqua_data))

aqua_data_monthly<-aqua_data |> 
  mutate(Date_Time_UTC_ = lubridate::ymd_hms(Date_Time_UTC_)) |> 
  mutate(month = floor_date(Date_Time_UTC_, unit = "month")) |> 
  group_by(Location_Name, Latitude, Longitude) |> 
  summarise(
    mean_temp = mean(Value, na.rm = TRUE),
    min_temp = min(Value, na.rm = TRUE),
    max_temp = max(Value, na.rm = TRUE),
    n_readings = sum(!is.na(Value)),
    min_date = min(Date_Time_UTC_, na.rm = TRUE),
    max_date = max(Date_Time_UTC_, na.rm = TRUE),
    .groups = "drop"
  )

### Filter out the location with temperatures above 50 degrees average
aqua_data_monthly <- aqua_data_monthly |> 
  filter(mean_temp < 50)

aqua_monthly_sf<-sf::st_as_sf(aqua_data_monthly, coords = c("Longitude", "Latitude"), crs = 4326)
  

ggplot()+ 
  geom_sf(data = bc) + 
  geom_sf(data = aqua_monthly_sf, aes(color = mean_temp)) +
  scale_color_viridis_c() +
  labs(title = "Mean Water Temperature from BC Aquarius Data", color = "Mean Temp (°C)") +
  theme_minimal()

## now a leaflet map, showing the locations and the mean, max, min temperatures and the number of readings
leaflet(aqua_monthly_sf) |> 
  addTiles() |> 
  addCircleMarkers(
    radius = 2,  # You can replace this with a scaling function based on n_readings if desired
    color = ~colorNumeric("viridis", mean_temp)(mean_temp),
    fillOpacity = 0.7,
    popup = popupTable(
      aqua_monthly_sf,
      zcol = c("Location_Name", "mean_temp", "min_temp", "max_temp", "min_date", "max_date", "n_readings")
    )
  ) |> 
  addLegend(
    position = "bottomright",
    pal = colorNumeric("viridis", aqua_monthly_sf$mean_temp),
    values = aqua_monthly_sf$mean_temp,
    title = "Mean Temperature (°C)"
  )

aqua_data_monthly |> 
  DT::datatable(
    options = list(pageLength = 10, autoWidth = TRUE),
    rownames = FALSE,
    filter = "top"
  )

```



#### Fish Occurrence Records
```{r}
# Fish occurrence records queried from typical data sources.
all_occs = sf::read_sf(paste0(lan_root,"2 SCIENCE - Invasives/SPECIES/5_Incidental Observations/AIS_occurrences_all_sources.gpkg"))
source("grab_occs_updated_path.R")
# rel_occs = all_occs |> 
#   dplyr::filter(Species %in% c("Bull trout","Sockeye salmon"))

# Add queries for Bull Trout and Sockeye Salmon
if(requery_occs){
  bt_occs = suppressMessages(suppressWarnings(grab_aq_occ_data_path_update('Bull Trout', quiet = T)))
  ss_occs = suppressMessages(suppressWarnings(grab_aq_occ_data_path_update('Sockeye Salmon')))
  
  sf::write_sf(bt_occs, "data/bull_trout_occurrences.gpkg")
  sf::write_sf(ss_occs, "data/sockeye_salmon_occurrences.gpkg")
} else {
  bt_occs = sf::read_sf("data/bull_trout_occurrences.gpkg")
  ss_occs = sf::read_sf("data/sockeye_salmon_occurrences.gpkg")
}

fish_occs = dplyr::bind_rows(bt_occs, ss_occs)

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = fish_occs) +
  facet_wrap( ~ Species)
```

#### Angler Use
This is the DFO data layer of angler use, with the scale being the number of days fished.
```{r plot_angler_use}

anglerUse<-terra::rast(paste0(onedrive_wd,"CNF/DFO_angling_survey_days_fished_raster.tif"))
#plot(anglerUse, main = "Angler Use (Days Fished)")

ggplot()+
  geom_sf(data = bc) + 
  geom_raster(data = as.data.frame(anglerUse, xy = TRUE), aes(x = x, y = y, fill = days_fished)) +
  scale_fill_viridis_c() +
  labs(title = "Angler Use (Days Fished)", fill = "Days Fished") +
  theme_minimal()

```


#### Aquaculture Licences

Licensing information is available from https://open.canada.ca/data/en/dataset/522d1b67-30d8-4a34-9b62-5da99b1035e6

```{r aquaculture_licences}

landbased<-read.csv("data/aquaculture licensing/licence-permis-landbased-terre-rpt-pac-dfo-mpo-aquaculture-eng.csv")

marine<-read.csv("data/aquaculture licensing/license-permis-mer-mar-rpt-pac-dfo-mpo-aquaculture-eng.csv")

enchancements<-read.csv("data/aquaculture licensing/licence-permis-enhancement-renforcement-rpt-pac-dfo-mpo-aquaculture-eng.csv")

landbased_salmonids<-landbased |> 
  filter(grepl("Salmon|trout|whitefish", Licensed.Species, ignore.case = TRUE))

marine_salmonids<-marine |>
  filter(grepl("Salmon|trout|whitefish", Licensed.Species, ignore.case = TRUE))

enchancements_salmonids<-enchancements |>
  filter(grepl("Salmon|trout|whitefish", Licensed.Species, ignore.case = TRUE))

leaflet() |>
  addTiles() |>

  # Land-based salmonids
  addCircleMarkers(
    data = sf::st_as_sf(landbased_salmonids, coords = c("Longitude", "Latitude"), crs = 4326),
    color = "blue",
    fillOpacity = 0.7,
    radius = 6,
    popup = lapply(
      popupTable(landbased_salmonids |> sf::st_drop_geometry()),
      HTML
    )
  ) |>

  # Marine salmonids
  addCircleMarkers(
    data = sf::st_as_sf(marine_salmonids, coords = c("Longitude", "Latitude"), crs = 4326),
    color = "green",
    fillOpacity = 0.7,
    radius = 6,
    popup = lapply(
      popupTable(marine_salmonids |> sf::st_drop_geometry()),
      HTML
    )
  ) |>

  # Enhancement salmonids
  addCircleMarkers(
    data = sf::st_as_sf(enchancements_salmonids, coords = c("Longitude", "Latitude"), crs = 4326),
    color = "red",
    fillOpacity = 0.7,
    radius = 6,
    popup = lapply(
      popupTable(enchancements_salmonids |> sf::st_drop_geometry()),
      HTML
    )
  ) |> 
  addLegend(
    position = "bottomleft",
    colors = c("blue", "green", "red"),
    labels = c("Land-based Salmonids", "Marine Salmonids", "Enhancements Salmonids"),
    title = "Salmonid Types",
    opacity = 0.7
  )

```

We also found a Pacific Salmon Foundation ArcGIS Online map that shows a bunch of hatchery locations; We cannot download the data, though.

[Link to PSF map](https://www.arcgis.com/apps/mapviewer/index.html?webmap=4b89fba48fa14acf8ef2f913bee76144)


#### Boat Launches

```{r}
ramps = readr::read_csv("data/FFSBC_RAMPS_Database.csv")

# Data clean-up
ramps_sf = ramps |> 
  dplyr::filter(!is.na(X),!is.na(Y)) |> 
  # Correct one typo found in the datat
  dplyr::mutate(Y = ifelse(WBID == "00542UDEN", 5832173, Y)) |> 
  # Remove some troublesome character encodings in 'Nearest_Town' field
  dplyr::mutate(Nearest_Town = stringr::str_remove_all(Nearest_Town,"(\xe8,\\?)")) |> 
  dplyr::group_by(UTM_Zone) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    unique_utm_zone = unique(.x$UTM_Zone)
    utm_zone = dplyr::case_when(
      unique_utm_zone == 8 ~ 32608,
      unique_utm_zone == 9 ~ 32609,
      unique_utm_zone == 10 ~ 32610,
      unique_utm_zone == 11 ~ 32611
    )
    .x |> 
      sf::st_as_sf(coords = c("X","Y"),
                   crs = utm_zone) |> 
      sf::st_transform(4326)
  }) |> 
  dplyr::bind_rows()

boats_sf = ramps_sf |>
  dplyr::filter(Boat_Launches != "NO")

popup_tbls = leafpop::popupTable(
        boats_sf |> 
          sf::st_drop_geometry() |> 
          dplyr::select(
            WBID,
            Gazetted_Name,
            Nearest_Town,
            Source_Lake_Parameters,
            Area_Surface_ha,
            Depth_Max_m,
            Depth_Mean_m,
            Boat_Launches,
            Access_Info_Source,
            AvgOfpH
          )
      )

leaflet() |> 
  addTiles() |> 
  addCircleMarkers(
    data = boats_sf,
    label = ~dplyr::coalesce(Gazetted_Name,WBID),
    popup = lapply(
      popup_tbls,
      htmltools::HTML
    )
  )

DT::datatable(boats_sf |> 
                dplyr::select(-c(UTM_Zone,Lat,Longitude)) |> 
                sf::st_drop_geometry())
```

#### Watercraft Inspections

These data contain the 

```{r water_craft_inspections}

library(patchwork)
#watercraftInspections<-sf::read_sf("W:/CMadsen/shared_data_sets/Waterbodies_with_Inspection_Data_Summaries.gpkg")

watercraft2<-sf::read_sf("W:/CMadsen/Projects/ZQMussels/data/Waterbodies_with_Inspection_Data_Summaries_all_years.gpkg")

p1<-ggplot()+
  geom_sf(data = bc) + 
  geom_sf(data = watercraft2, aes(color = TotalInspections)) +
  scale_color_viridis_c() +
  labs(title = "Watercraft Inspections", color = "Inspection Count") +
  theme_minimal()

p2<-ggplot()+
  geom_sf(data = bc) + 
  geom_sf(data = watercraft2, aes(color = highrisk_Counter)) +
  scale_color_viridis_c() +
  labs(title = "Watercraft Inspections", color = "No. high risk watercraft") +
  theme_minimal()

p1

p2


```


#### Distance to roads

The distance to numbered highways is a raster layer that shows the distance to the nearest numbered highway in meters. 
```{r plot_road_distances}

road_distances<-terra::rast(paste0(onedrive_wd,"CNF/distance_to_numbered_highway_raster.tif"))

ggplot()+
  geom_sf(data = bc) + 
  geom_raster(data = as.data.frame(road_distances, xy = TRUE), aes(x = x, y = y, fill = layer)) +
  scale_fill_viridis_c() +
  labs(title = "Distance to Numbered Highways", fill = "Distance (m)") +
  theme_minimal()
```



#### Fish Periodicity

I'm just waiting on Amy Sigvaldson's work to amalgamate edits from Ron Ptolemy, DFO, and other groups to finalize the fish periodicity file. This file describes when species' life stages occur in major rivers.

Amy's data is at: `\\sfp.idir.bcgov\s140\S40203\WFC AEB\General\2 Stewardship Mgt Climate Change\Drought\Communications\Periodicity tables`



