---
title: "Effect of bins"
author: "John Phelan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(readxl)
library(BAMMtools)
library(terra)
#devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)
library(tidyterra)
library(sf)

options(scipen = 999)

k_breaks = 3

base_dir = stringr::str_extract(getwd(),"C:\\/Users\\/[a-zA-Z]+")
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")
lan_root = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/"
# proj_wd = getwd()
# if(stringr::str_detect(proj_wd,"scripts")){
#   proj_wd = stringr::str_remove(proj_wd, "scripts/markdown")
# }


# This function finds the natural breaks (or, "Jenks") for a numeric column
bin_to_natural_breaks = function(dat,variable){
  # find natural breaks ("jenks")
  nbreaks = BAMMtools::getJenksBreaks(var = dat[[variable]], k = 4)
  dat |> 
    dplyr::rename(var_to_bin = !!rlang::sym(variable)) |> 
    dplyr::mutate(var_bin = as.numeric(cut(var_to_bin,nbreaks))) |> 
    dplyr::mutate(var_bin = tidyr::replace_na(var_bin, 1)) |> 
    dplyr::mutate(var_bin = as.character(var_bin)) |> 
    dplyr::mutate(var_bin = factor(var_bin, levels = c(1:3))) |> 
    dplyr::rename(!!rlang::sym(paste0(variable,"_natural_bin")) := var_bin,
                  !!rlang::sym(variable) := var_to_bin)
}

#k-means clustering function
bin_to_kmeans <- function(dat, variable, centers = k_breaks) {
  has_geometry <- inherits(dat, "sf")
  
  # Store and drop geometry if present
  if (has_geometry) {
    geometry <- sf::st_geometry(dat)
    dat_df <- sf::st_drop_geometry(dat)
  } else {
    dat_df <- dat
  }

  # Identify valid rows
  valid_rows <- is.finite(dat_df[[variable]])
  dat_df_valid <- dat_df[valid_rows, ]

  # Run k-means clustering
  set.seed(123)
  km <- stats::kmeans(dat_df_valid[[variable]], centers = centers, iter.max = 100)

  # Reorder cluster labels by mean value
  cluster_means <- tapply(dat_df_valid[[variable]], km$cluster, mean)
  ordered_clusters <- order(cluster_means)
  relabeled_clusters <- match(km$cluster, ordered_clusters)

  # Assign bins back to full dataset
  dat_df$var_bin <- NA
  dat_df$var_bin[valid_rows] <- relabeled_clusters

  # Final formatting
  dat_df <- dat_df |> 
    dplyr::rename(var_to_bin = !!rlang::sym(variable)) |> 
    dplyr::mutate(var_bin = as.character(var_bin)) |> 
    dplyr::mutate(var_bin = factor(var_bin, levels = as.character(1:centers))) |> 
    dplyr::rename(
      !!rlang::sym(paste0(variable, "_kmeans_bin")) := var_bin,
      !!rlang::sym(variable) := var_to_bin
    )

  # Reattach geometry if present
  if (has_geometry) {
    dat_df <- sf::st_as_sf(dat_df)
    sf::st_geometry(dat_df) <- geometry
  }

  return(dat_df)
}



## GGplot label convenience function
bin_info_labs = function(bins_title){
    labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)',
       subtitle = paste0("Inspection Bins: 1 ('low', ",
                         bins_title[1]," to ",bins_title[2],
                         "), 2 ('medium', ",
                         bins_title[2]," to ",bins_title[3],
                         "), 3 ('high', ",
                         bins_title[3]," to ",bins_title[4],
                         ")"))
}

getKMeansBreaks <- function(var, k = 3) {
  var_clean <- var[!is.na(var)]  # remove NAs
  set.seed(123)  # for reproducibility
  km <- stats::kmeans(var_clean, centers = k)
  centers <- sort(km$centers)
  # Compute midpoints between sorted centers
  breaks <- c(min(var_clean),
              zoo::rollmean(centers, 2),
              max(var_clean))
  return(breaks)
}

```

```{r data}

insp_from_wd_inf_to_wb = sf::read_sf("W:/CMadsen/Projects/ZQMussels/data/Waterbodies_with_Inspection_Data_Summaries_all_years_WD_Infected_Areas.gpkg")
insp_from_wd_bc_wb = sf::read_sf("W:/CMadsen/Projects/ZQMussels/data/Waterbodies_with_Inspection_Data_Summaries_all_years_BC_WD_Infected_Waterbodies.gpkg")
ang = sf::read_sf("W:/CMadsen/shared_data_sets/freshwater_fisheries_society_angler_survey_2022_2023.gpkg")


```

```{r binning_inspection_data}

insp_from_wd_inf_to_wb_b = bin_to_natural_breaks(insp_from_wd_inf_to_wb, 'TotalInspections')

insp_from_wd_inf_to_wb_b = bin_to_kmeans(insp_from_wd_inf_to_wb_b, 'TotalInspections')

insp_from_wd_inf_to_wb_b = insp_from_wd_inf_to_wb_b |> 
  dplyr::filter(!GNIS_NA %in% c("Pacific Ocean","Dry Storage"))

```

```{r binning_inspection_data_bc}

insp_from_wd_bc_wb_b = bin_to_natural_breaks(insp_from_wd_bc_wb, 'TotalInspections')
insp_from_wd_bc_wb_b = bin_to_kmeans(insp_from_wd_bc_wb_b, 'TotalInspections')

insp_from_wd_bc_wb_b = insp_from_wd_bc_wb_b |> 
  dplyr::filter(!GNIS_NA %in% c("Pacific Ocean","Dry Storage"))

```

```{r binning_angler_data}
ang_b = bin_to_natural_breaks(ang, 'days_fished')
ang_b = bin_to_kmeans(ang_b, 'days_fished')

```



## Binning data {.tabset}

Natural breaks identify groupings by minimizing variance within bins and maximizing variance between bins based on the data distribution. Essentially, this method finds “natural” gaps or clusters in the data and sets bin boundaries accordingly. It is especially useful when the data are unevenly distributed or contain distinct clusters, as it adapts to the inherent structure of the data.

K-means clustering is a method that partitions data into a specified number of groups by minimizing the distance between data points and their assigned cluster centers. It tries to create clusters with similar values by iteratively adjusting cluster centers. The bin boundaries are then derived from these cluster centers. This approach assumes roughly spherical clusters and tends to balance cluster sizes, which can lead to different bin cutoffs compared to natural breaks.

The natural breaks method produced bins that align with natural groupings in the  data, capturing irregular clusters and gaps. In contrast, k-means bin boundaries are influenced by the overall grouping of data points into clusters, potentially smoothing over irregularities.

Choosing between these methods depends on the goal. If the intent is to highlight distinct groupings in effort that reflect natural patterns, natural breaks may be better. If the aim is to create balanced group sizes or identify typical cluster centers, k-means may be more appropriate.


### Natural breaks - Inspections from out of province

```{r break_values_insp}
bins_title = BAMMtools::getJenksBreaks(var = insp_from_wd_inf_to_wb$TotalInspections, k = 4)

bin_assignments <- cut(
  insp_from_wd_inf_to_wb$TotalInspections,
  breaks = bins_title,
  include.lowest = TRUE,
  labels = FALSE
)

```

The inspection bins are: 1 ('low', `r bins_title[1]` to `r bins_title[2]`), 2 ('medium', `r bins_title[2]` to `r bins_title[3]`), and 3 ('high', `r bins_title[3]` to `r bins_title[4]`).


```{r natural_breaks}
# plots of the natural break data


nat_inspec<-ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = insp_from_wd_inf_to_wb_b, aes(fill = TotalInspections_natural_bin,
                                               col = TotalInspections_natural_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)')

nat_inspec

```

### Natural breaks - Inspections from BC infected waterbodies

```{r break_value_insp_bc}
bins_title_bc = BAMMtools::getJenksBreaks(var = insp_from_wd_bc_wb$TotalInspections, k = 4)

```

The inspection bins are: 1 ('low', `r bins_title_bc[1]` to `r bins_title_bc[2]`), 2 ('medium', `r bins_title_bc[2]` to `r bins_title_bc[3]`), and 3 ('high', `r bins_title_bc[3]` to `r bins_title_bc[4]`).

```{r plot_inspections_bc}

nat_inspec_bc<-ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = insp_from_wd_bc_wb_b, aes(fill = TotalInspections_natural_bin,
                                               col = TotalInspections_natural_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)')
nat_inspec_bc
```

### Natural breaks - Angler Data

```{r break_value_angler}
bins_title_ang = BAMMtools::getJenksBreaks(var = ang$days_fished, k = 4)
```

The angler bins are: 1 ('low', `r bins_title_ang[1]` to `r bins_title_ang[2]`), 2 ('medium', `r bins_title_ang[2]` to `r bins_title_ang[3]`), and 3 ('high', `r bins_title_ang[3]` to `r bins_title_ang[4]`).

```{r plot_angler}
nat_angler<-ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = ang_b, aes(fill = days_fished_natural_bin,
                                               col = days_fished_natural_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Days Fished (bin)',
       fill = 'Days Fished (bin)')

nat_angler
```

### K-means - Inspections from out of province

```{r kmeans_inspection}


bins_kmeans <- getKMeansBreaks(insp_from_wd_inf_to_wb$TotalInspections, k_breaks)

```

The inspection bins are: 1 ('low', `r round(bins_kmeans[1],0)` to `r round(bins_kmeans[2],0)`), 2 ('medium', `r round(bins_kmeans[2],0)` to `r round(bins_kmeans[3],0)`), and 3 ('high', `r round(bins_kmeans[3],0)` to `r round(bins_kmeans[4],0)`).

```{r kmeans_inspection_plot}
kmeans_inspec<-ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = insp_from_wd_inf_to_wb_b, aes(fill = TotalInspections_kmeans_bin,
                                               col = TotalInspections_kmeans_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)')
kmeans_inspec
```

### K-means - Inspections from BC infected waterbodies
```{r kmeans_inspection_bc}
bins_kmeans_bc <- getKMeansBreaks(insp_from_wd_bc_wb$TotalInspections, k_breaks)
```

The inspection bins are: 1 ('low', `r round(bins_kmeans_bc[1],0)` to `r round(bins_kmeans_bc[2],0)`), 2 ('medium', `r round(bins_kmeans_bc[2],0)` to `r round(bins_kmeans_bc[3],0)`), and 3 ('high', `r round(bins_kmeans_bc[3],0)` to `r round(bins_kmeans_bc[4],0)`).

```{r kmeans_inspection_bc_plot}
kmeans_inspec_bc<-ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = insp_from_wd_bc_wb_b, aes(fill = TotalInspections_kmeans_bin,
                                               col = TotalInspections_kmeans_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)')
kmeans_inspec_bc
```

### K-means - Angler Data
```{r kmeans_angler}
bins_kmeans_ang <- getKMeansBreaks(ang$days_fished, k_breaks)
```

The angler bins are: 1 ('low', `r round(bins_kmeans_ang[1],0)` to `r round(bins_kmeans_ang[2],0)`), 2 ('medium', `r round(bins_kmeans_ang[2],0)` to `r round(bins_kmeans_ang[3],0)`), and 3 ('high', `r round(bins_kmeans_ang[3],0)` to `r round(bins_kmeans_ang[4],0)`).

```{r kmeans_angler_plot}
kmeans_angler<-ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = ang_b, aes(fill = days_fished_kmeans_bin,
                                               col = days_fished_kmeans_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Days Fished (bin)',
       fill = 'Days Fished (bin)')
kmeans_angler
```

### Comparison of Binning Methods
```{r comparison_of_bins, fig.width = 12, fig.height = 8}
gridExtra::grid.arrange(nat_inspec, kmeans_inspec, nat_inspec_bc, kmeans_inspec_bc, nat_angler, kmeans_angler,
                         ncol = 2, nrow = 3,
                         top = "Comparison of Binning Methods for Inspections and Angler Data")
```

### Summary of Binning Methods
```{r summary_of_bins}
summary_bins <- tibble(
  Method = c("Natural Breaks", "K-means"),
  Inspections_Out_of_Province = c(
    paste0("1: ", round(bins_title[1]), " to ", round(bins_title[2]), 
           ", 2: ", round(bins_title[2]), " to ", round(bins_title[3]), 
           ", 3: ", round(bins_title[3]), " to ", round(bins_title[4])),
    paste0("1: ", round(bins_kmeans[1]), " to ", round(bins_kmeans[2]), 
           ", 2: ", round(bins_kmeans[2]), " to ", round(bins_kmeans[3]), 
           ", 3: ", round(bins_kmeans[3]), " to ", round(bins_kmeans[4]))
  ),
  Inspections_BC_Infected = c(
    paste0("1: ", round(bins_title_bc[1]), " to ", round(bins_title_bc[2]), 
           ", 2: ", round(bins_title_bc[2]), " to ", round(bins_title_bc[3]), 
           ", 3: ", round(bins_title_bc[3]), " to ", round(bins_title_bc[4])),
    paste0("1: ", round(bins_kmeans_bc[1]), " to ", round(bins_kmeans_bc[2]), 
           ", 2: ", round(bins_kmeans_bc[2]), " to ", round(bins_kmeans_bc[3]), 
           ", 3: ", round(bins_kmeans_bc[3]), " to ", round(bins_kmeans_bc[4]))
  ),
  Angler_Days_Fished = c(
    paste0("1: ", round(bins_title_ang[1]), " to ", round(bins_title_ang[2]), 
           ", 2: ", round(bins_title_ang[2]), " to ", round(bins_title_ang[3]), 
           ", 3: ", round(bins_title_ang[3]), " to ", round(bins_title_ang[4])),
    paste0("1: ", round(bins_kmeans_ang[1]), " to ", round(bins_kmeans_ang[2]), 
           ", 2: ", round(bins_kmeans_ang[2]), " to ", round(bins_kmeans_ang[3]), 
           ", 3: ", round(bins_kmeans_ang[3]), " to ", round(bins_kmeans_ang[4]))
  )
)

summary_bins |> 
  knitr::kable(caption = "Summary of Binning Methods for Inspections and Angler Data")
```

### Comparison of Inspection Data Binning Methods

```{r insp_comparison}
insp_b_diff <- insp_from_wd_inf_to_wb_b |> 
  st_drop_geometry() |> 
  mutate(
    natural_rank = as.numeric(factor(TotalInspections_natural_bin, ordered = TRUE)),
    kmeans_rank = as.numeric(factor(TotalInspections_kmeans_bin, ordered = TRUE)),
    rank_diff = kmeans_rank - natural_rank) |> 
  arrange(desc(abs(rank_diff))) |>  # most different at top
  dplyr::select(
    WATERSH,
    GNIS_NA,
    TotalInspections_natural_bin,
    TotalInspections_kmeans_bin,
    natural_rank,
    kmeans_rank,
    rank_diff
  )
# View changes
DT::datatable(insp_b_diff, 
               options = list(pageLength = 10, 
                              autoWidth = TRUE, 
                              columnDefs = list(list(className = 'dt-center', targets = "_all"))),
               rownames = FALSE,
               caption = "Comparison of Inspection Data Binning Methods")
```

### Comparison of Inspection Data Binning Methods (BC Infected)
```{r insp_comparison_bc}
insp_b_diff_bc <- insp_from_wd_bc_wb_b |> 
  st_drop_geometry() |> 
  mutate(
    natural_rank = as.numeric(factor(TotalInspections_natural_bin, ordered = TRUE)),
    kmeans_rank = as.numeric(factor(TotalInspections_kmeans_bin, ordered = TRUE)),
    rank_diff = kmeans_rank - natural_rank) |> 
  arrange(desc(abs(rank_diff))) |>   # most different at top 
  dplyr::select(
    WATERSH,
    GNIS_NA,
    TotalInspections_natural_bin,
    TotalInspections_kmeans_bin,
    natural_rank,
    kmeans_rank,
    rank_diff
  )
# View changes
DT::datatable(insp_b_diff_bc, 
               options = list(pageLength = 10, 
                              autoWidth = TRUE, 
                              columnDefs = list(list(className = 'dt-center', targets = "_all"))),
               rownames = FALSE,
               caption = "Comparison of Inspection Data Binning Methods (BC Infected)")
```


### Comparison of Angler Data Binning Methods

```{r ang_comparison}
ang_b_diff <- ang_b |> 
  st_drop_geometry() |> 
  mutate(
    natural_rank = as.numeric(factor(days_fished_natural_bin, ordered = TRUE)),
    kmeans_rank = as.numeric(factor(days_fished_kmeans_bin, ordered = TRUE)),
    rank_diff = kmeans_rank - natural_rank) |> 
  arrange(desc(abs(rank_diff))) |>  
  dplyr::select(
    WATERSHED_GROUP_ID,
    Waterbody,
    days_fished_natural_bin,
    days_fished_kmeans_bin,
    natural_rank,
    kmeans_rank,
    rank_diff
  )
  
# View changes
DT::datatable(
  ang_b_diff |> dplyr::rename(WATERSH = WATERSHED_GROUP_ID),
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames = FALSE,
  caption = "Comparison of Angler Data Binning Methods"
)

```

### Priority Scores - Difference in method scores
```{r priority_scores_natural_breaks}

# 1. Drop geometry and select relevant columns with native pipe

inf_to_bc_bins <- insp_from_wd_inf_to_wb_b |>
  sf::st_drop_geometry() |>
  dplyr::select(
    WATERSH,
    GNIS_NA,
    inf_to_bc_nat_bin = TotalInspections_natural_bin,
    inf_to_bc_kmns_bin = TotalInspections_kmeans_bin
  )

bc_wd_bins <- insp_from_wd_bc_wb_b |>
  sf::st_drop_geometry() |>
  dplyr::select(
    WATERSH,
    GNIS_NA,
    bc_wd_nat_bin = TotalInspections_natural_bin,
    bc_wd_kmns_bin = TotalInspections_kmeans_bin
  )

ang_bins <- ang_b |>
  sf::st_drop_geometry() |>
  dplyr::rename(
    WATERSH = WATERSHED_GROUP_ID,
    GNIS_NA = Waterbody
  ) |>
  dplyr::select(
    WATERSH,
    GNIS_NA,
    ang_nat_bin = days_fished_natural_bin,
    ang_kmns_bin = days_fished_kmeans_bin
  )

# 2. Join and calculate separate priorities

priority_table <- inf_to_bc_bins |>
  dplyr::full_join(bc_wd_bins, by = c("WATERSH", "GNIS_NA")) |>
  dplyr::full_join(ang_bins, by = c("WATERSH", "GNIS_NA")) |>
  dplyr::mutate(
    # Convert factors/characters to numeric if needed
    inf_to_bc_nat_bin = as.numeric(as.character(inf_to_bc_nat_bin)),
    bc_wd_nat_bin = as.numeric(as.character(bc_wd_nat_bin)),
    ang_nat_bin = as.numeric(as.character(ang_nat_bin)),
    inf_to_bc_kmns_bin = as.numeric(as.character(inf_to_bc_kmns_bin)),
    bc_wd_kmns_bin = as.numeric(as.character(bc_wd_kmns_bin)),
    ang_kmns_bin = as.numeric(as.character(ang_kmns_bin))
  ) |>
  dplyr::mutate(
    # Replace NAs with 0
    inf_to_bc_nat_bin = tidyr::replace_na(inf_to_bc_nat_bin, 0),
    bc_wd_nat_bin = tidyr::replace_na(bc_wd_nat_bin, 0),
    ang_nat_bin = tidyr::replace_na(ang_nat_bin, 0),
    inf_to_bc_kmns_bin = tidyr::replace_na(inf_to_bc_kmns_bin, 0),
    bc_wd_kmns_bin = tidyr::replace_na(bc_wd_kmns_bin, 0),
    ang_kmns_bin = tidyr::replace_na(ang_kmns_bin, 0)
  ) |>
  dplyr::mutate(
    # Calculate separate total priorities
    total_priority_natural = inf_to_bc_nat_bin + bc_wd_nat_bin + ang_nat_bin,
    total_priority_kmeans = inf_to_bc_kmns_bin + bc_wd_kmns_bin + ang_kmns_bin,
    difference_methods = total_priority_natural - total_priority_kmeans
  )

DT::datatable(priority_table, 
               options = list(pageLength = 10, 
                              autoWidth = TRUE, 
                              columnDefs = list(list(className = 'dt-center', targets = "_all"))),
               rownames = FALSE,
               caption = "Priority Scores for Natural Breaks and K-means Binning Methods")
```



