---
title: "fish_temp_overlap"
author: "John Phelan"
date: "2025-05-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(terra)
library(dbplyr)
library(DBI)
library(tidyverse)
library(bcinvadeR)
library(bcdata)
library(RColorBrewer)
library(leaflet)
library(leafpop)

base_dir = stringr::str_extract(getwd(),"C:\\/Users\\/[a-zA-Z]+")
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")
lan_root = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/"

requery_occs = F

# Custom function to show number of rows with commas.
nrowc = function(dat){
  scales::comma(nrow(dat))
}
```


#### Hydroclim
```{r get_hydroclim}
hydroclim_layers_in_onedrive = list.files(
  path = paste0(onedrive_wd,"raster/hydroclim_goldMussel_rast/"),
  pattern = "^bc_.*\\.tif$",
  full.names = T
)

hydroclim_layer_names = paste0(str_remove(str_replace_all(str_extract(hydroclim_layers_in_onedrive,"bc_.*\\.tif$"),"(_|\\.)"," "),"( rast)? tif"), collapse = ', ')

hydroclim = terra::rast(hydroclim_layers_in_onedrive[1])


```

#### Occurrences

The species that are susseptible to whirling disease are: 
- Oncorhynchus clarkii (cutthroat trout)
- Oncorhynchus kisutch (coho salmon)
- Oncorhynchus mykiss (rainbow trout)
- Oncorhynchus nerka (sockeye salmon)
- Oncorhynchus tshawytscha (chinook salmon)
- Prosopium williamsoni (mountain whitefish)
- Salmo salar (Atlantic salmon)
- Salmo trutta (brown trout)
- Salvelinus confluentus (bull trout)
- Salvelinus fontinalis (brook trout)

https://inspection.canada.ca/en/animal-health/aquatic-animals/diseases/reportable-diseases/whirling-disease/fact-sheet


```{r fish_locations}
# Fish occurrence records queried from typical data sources.
all_occs = sf::read_sf(paste0(lan_root,"2 SCIENCE - Invasives/SPECIES/5_Incidental Observations/AIS_occurrences_all_sources.gpkg"))
source("grab_occs_updated_path.R")
# rel_occs = all_occs |> 
#   dplyr::filter(Species %in% c("Bull trout","Sockeye salmon"))


species_names <- c(
  "Bull Trout", "Sockeye Salmon", "Cutthroat Trout", "Coho Salmon",
  "Rainbow Trout", "Chinook Salmon", "Mountain Whitefish", 
  "Atlantic Salmon", "Brown Trout", "Brook Trout"
)
if(requery_occs){
species_occs <- lapply(species_names, function(sp) {
  ocs<-suppressMessages(suppressWarnings(grab_aq_occ_data_path_update(sp, quiet = TRUE)))
   if (!is.null(ocs)) {
    outfile <- paste0("data/", gsub(" ", "_", sp), "_occurrences.gpkg")
    sf::write_sf(ocs, outfile)
  }
  return(ocs)
})

}else{
  species_occs <- lapply(species_names, function(sp) {
    sf::read_sf(paste0("data/", gsub(" ", "_", sp), "_occurrences.gpkg"))
  })
}
names(species_occs) <- gsub(" ", "_", species_names)

bc = bcmaps::bc_bound() |> 
  sf::st_transform(4326)
fish_occs <- bind_rows(species_occs, .id = "species")

ggplot() + 
  geom_sf(data = bc, color = "lightgrey") + 
  geom_sf(data = fish_occs) +
  facet_wrap( ~ Species)+
  theme_minimal()
  
```


#### High Priority Species

The species with deemed to have the highest susceptability are widely distributed throughout the province

```{r high_priority_species}

high_priority_species <- c(
  "Sockeye Salmon", "Rainbow Trout", "Cutthroat Trout"
)

selected_sp <- fish_occs |> 
  filter(Species %in% high_priority_species)


ggplot() +
  geom_sf(data = bc, color = "lightgrey") + 
  geom_sf(data = selected_sp, aes (color = Species), size = 0.75) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap( ~ Species)+
  theme_minimal()+
  labs(title = "High Priority Fish Species Occurrences in BC",
       subtitle = "Species: Sockeye Salmon, Rainbow (Steelhead) Trout, Cutthroat Trout")



```

```{r minAirTemps}

# Find only the files with "Min.monthlyairtemperaturefor" in their filename
min_temp_files <- grep("Min\\.monthlyairtemperaturefor", hydroclim_layers_in_onedrive, value = TRUE)


# Extract clean layer names from filenames for use as list names
min_temp_names <- gsub("_rast\\.tif$", "", basename(min_temp_files))

# Load rasters into a named list
min_hydroclim <- lapply(seq_along(min_temp_files), function(i) {
  r <- rast(min_temp_files[i])
  names(r) <- min_temp_names[i]
  r
})
names(min_hydroclim) <- min_temp_names

```

#### Overlap of Hydroclim temperatures and Fish Occurrences

The hydroclim data is the average minimum temperature across the province, at 2km resolution. For each month, the tmperature where each fish occurrence is located is extracted, and then the fish occurrences are subsetted to only those that are greater than 10 degrees and less than 17 degrees.


```{r subset temps by spp occs}

# for each fish occurrence, get the montly temperature at that associated temperature.
# we will build a data frame for each point showing the location, month and temperature

# List of month names for ordering
month_names <- gsub("bc_Min.monthlyairtemperaturefor|_", "", names(min_hydroclim))
month_names <- gsub("_", "", month_names)

test<-selected_sp
selected_sp<-test
if (inherits(selected_sp, "sf")) {
  selected_sp_vect <- terra::vect(selected_sp)  # this preserves geometry
} else {
  stop("selected_sp must be an sf object")
}

# Load raster stack
min_hydroclim_stack <- terra::rast(min_hydroclim)

# Extract raster values
extracted_values <- terra::extract(min_hydroclim_stack, selected_sp_vect)

# Add fish_id for merging
selected_sp$fish_id <- 1:nrow(selected_sp)

# Add lon/lat columns to sf object for later plotting
selected_sp$lon <- sf::st_coordinates(selected_sp)[, 1]
selected_sp$lat <- sf::st_coordinates(selected_sp)[, 2]

# Merge the extracted temperature values with the spatial data
fish_temps_df_wide <- merge(
  as.data.frame(selected_sp),
  extracted_values,
  by.x = "fish_id",
  by.y = "ID",
  all.x = TRUE
)

# Rename the monthly temperature columns
names(fish_temps_df_wide)[grep("bc_Min.monthlyairtemperaturefor", names(fish_temps_df_wide))] <- month_names

# Pivot to long format
fish_temps_df_long <- pivot_longer(
  fish_temps_df_wide,
  cols = all_of(month_names),
  names_to = "month",
  values_to = "temperature"
)

# Filter by temperature range
fish_temps_df_long <- fish_temps_df_long %>%
  filter(temperature >= 10, temperature <= 17)

# Convert to sf using saved lon/lat
fish_temps_df_long <- sf::st_as_sf(fish_temps_df_long, coords = c("lon", "lat"), crs = 4326)

# Set month factor levels for plotting
fish_temps_df_long$month <- factor(
  fish_temps_df_long$month,
  levels = c("January", "February", "March", "April", "May", "June",
             "July", "August", "September", "October", "November", "December")
)

# Plot
ggplot() +
  geom_sf(data = bc, color = "lightgrey") +
  geom_sf(data = fish_temps_df_long, aes(color = Species), size = 0.75) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~ month) +
  theme_minimal() +
  labs(
    title = "Fish Occurrences with Temperatures Between 10°C and 17°C",
    subtitle = "Faceted by Month"
  )+
   guides(color = guide_legend(override.aes = list(size = 2))) 


```


#### Filtering Aquarius temperature data





