---
title: "fish_temp_overlap"
author: "John Phelan & Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(terra)
library(dbplyr)
library(DBI)
library(tidyverse)
library(bcinvadeR)
library(bcdata)
library(RColorBrewer)
library(leaflet)
library(leafpop)
library(readxl)
library(sf)

base_dir = stringr::str_extract(getwd(),"C:\\/Users\\/[a-zA-Z]+")
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")
lan_root = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/"

requery_occs = F

# Custom function to show number of rows with commas.
nrowc = function(dat){
  scales::comma(nrow(dat))
}
```


### Overlap of Fish and Parameters {.tabset}

#### Hydroclim
```{r get_hydroclim}
hydroclim_layers_in_onedrive = list.files(
  path = paste0(onedrive_wd,"raster/hydroclim_goldMussel_rast/"),
  pattern = "^bc_.*\\.tif$",
  full.names = T
)

hydroclim_layer_names = paste0(str_remove(str_replace_all(str_extract(hydroclim_layers_in_onedrive,"bc_.*\\.tif$"),"(_|\\.)"," "),"( rast)? tif"), collapse = ', ')

hydroclim = terra::rast(hydroclim_layers_in_onedrive[1])


```

#### Occurrences

The species that are susseptible to whirling disease are: 
- Oncorhynchus clarkii (cutthroat trout)
- Oncorhynchus kisutch (coho salmon)
- Oncorhynchus mykiss (rainbow trout)
- Oncorhynchus nerka (sockeye salmon)
- Oncorhynchus tshawytscha (chinook salmon)
- Prosopium williamsoni (mountain whitefish)
- Salmo salar (Atlantic salmon)
- Salmo trutta (brown trout)
- Salvelinus confluentus (bull trout)
- Salvelinus fontinalis (brook trout)

https://inspection.canada.ca/en/animal-health/aquatic-animals/diseases/reportable-diseases/whirling-disease/fact-sheet


```{r fish_locations}
# Fish occurrence records queried from typical data sources.
all_occs = sf::read_sf(paste0(lan_root,"2 SCIENCE - Invasives/SPECIES/5_Incidental Observations/AIS_occurrences_all_sources.gpkg"))
source("grab_occs_updated_path.R")
# rel_occs = all_occs |> 
#   dplyr::filter(Species %in% c("Bull trout","Sockeye salmon"))


species_names <- c(
  "Bull Trout", "Sockeye Salmon", "Cutthroat Trout", "Coho Salmon",
  "Rainbow Trout", "Chinook Salmon", "Mountain Whitefish", 
  "Atlantic Salmon", "Brown Trout", "Brook Trout"
)
if(requery_occs){
species_occs <- lapply(species_names, function(sp) {
  ocs<-suppressMessages(suppressWarnings(grab_aq_occ_data_path_update(sp, quiet = TRUE)))
   if (!is.null(ocs)) {
    outfile <- paste0("data/", gsub(" ", "_", sp), "_occurrences.gpkg")
    sf::write_sf(ocs, outfile)
  }
  return(ocs)
})

}else{
  species_occs <- lapply(species_names, function(sp) {
    sf::read_sf(paste0("data/", gsub(" ", "_", sp), "_occurrences.gpkg"))
  })
}
names(species_occs) <- gsub(" ", "_", species_names)

bc = bcmaps::bc_bound() |> 
  sf::st_transform(4326)

fish_occs <- bind_rows(species_occs, .id = "species")

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = fish_occs, aes(color = as.factor(Species)))+
  facet_wrap( ~ Species)+
  scale_color_viridis_d() +
  theme_minimal()
  
```



```{r minAirTemps}

# Find only the files with "Min.monthlyairtemperaturefor" in their filename
min_temp_files <- grep("Min\\.monthlyairtemperaturefor", hydroclim_layers_in_onedrive, value = TRUE)


# Extract clean layer names from filenames for use as list names
min_temp_names <- gsub("_rast\\.tif$", "", basename(min_temp_files))

# Load rasters into a named list
min_hydroclim <- lapply(seq_along(min_temp_files), function(i) {
  r <- rast(min_temp_files[i])
  names(r) <- min_temp_names[i]
  r
})
names(min_hydroclim) <- min_temp_names

```

#### Overlap of Hydroclim temperatures and Fish Occurrences

The air temperature data ("Hydroclim") is the average minimum temperature across the province, at 2km resolution. For each month, the temperature at each fish occurrence is extracted and fish occurrences are filtered to only those with air temperature greater than 10 degrees Celsius.

```{r subset temps by spp occs}

# for each fish occurrence, get the montly temperature at that associated temperature.
# we will build a data frame for each point showing the location, month and temperature

# List of month names for ordering
month_names <- gsub("bc_Min.monthlyairtemperaturefor|_", "", names(min_hydroclim))
month_names <- gsub("_", "", month_names)

# test<-selected_sp
# selected_sp<-test
if (inherits(fish_occs, "sf")) {
  selected_sp_vect <- terra::vect(fish_occs)  # this preserves geometry
} else {
  stop("fish_occs must be an sf object")
}

# Load raster stack
min_hydroclim_stack <- terra::rast(min_hydroclim)

# Extract raster values
extracted_values <- terra::extract(min_hydroclim_stack, selected_sp_vect)

# Add fish_id for merging
fish_occs$fish_id <- 1:nrow(fish_occs)

# Add lon/lat columns to sf object for later plotting
fish_occs$lon <- sf::st_coordinates(fish_occs)[, 1]
fish_occs$lat <- sf::st_coordinates(fish_occs)[, 2]

# Merge the extracted temperature values with the spatial data
fish_temps_df_wide <- merge(
  as.data.frame(fish_occs),
  extracted_values,
  by.x = "fish_id",
  by.y = "ID",
  all.x = TRUE
)

# Rename the monthly temperature columns
names(fish_temps_df_wide)[grep("bc_Min.monthlyairtemperaturefor", names(fish_temps_df_wide))] <- month_names

# Pivot to long format
fish_temps_df_long <- pivot_longer(
  fish_temps_df_wide,
  cols = all_of(month_names),
  names_to = "month",
  values_to = "temperature"
)

# Filter by temperature range
fish_temps_df_long <- fish_temps_df_long |> 
  # filter(temperature >= 10, temperature <= 17)
  filter(temperature >= 10)

# Convert to sf using saved lon/lat
fish_temps_df_long <- sf::st_as_sf(fish_temps_df_long, coords = c("lon", "lat"), crs = 4326)

# Set month factor levels for plotting
fish_temps_df_long$month <- factor(
  fish_temps_df_long$month,
  levels = c("January", "February", "March", "April", "May", "June",
             "July", "August", "September", "October", "November", "December")
)

# Plot
ggplot() +
  geom_sf(data = bc, color = "lightgrey") +
  geom_sf(data = fish_temps_df_long, aes(color = Species), size = 0.75) +
  scale_color_viridis_d() +
  facet_wrap(~ month) +
  theme_minimal() +
  labs(
    # title = "Fish Occurrences with Temperatures Between 10째C and 17째C"
    title = "Fish Occurrences with Air Temperatures Above 10째C"
  )+
   guides(color = guide_legend(override.aes = list(size = 2))) 


```

#### Positive locations

Link to Whirling Disease Dashboard: https://bcgov-env.shinyapps.io/BCWhirlingDiseaseDashboard
```{r get_WD_results}

lan_folder = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/"
proj_wd = getwd()
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")

# Data
# Copy excel results from LAN folder IF it has all the right columns. This is
# intended to save us from overwriting the local data with some garbage
# LAN file that has replaced our goldenboy data file.
new_potential_dat_file = read_excel(paste0(lan_folder,"2 SCIENCE - Invasives/SPECIES/Whirling Disease/Monitoring/WD_sampling_results_fish_eDNA_used_for_making_maps_CMADSEN.xlsx"), sheet = "Fish and eDNA")

local_data_file = read_excel('data/WD_sampling_results_fish_eDNA_used_for_making_maps_CMADSEN.xlsx')

if(identical(names(new_potential_dat_file),names(local_data_file)) & nrow(new_potential_dat_file) == nrow(local_data_file)){
  print("New data file has identical column names and number of rows. Copying locally...")
  file.copy(from = paste0(lan_folder,"2 SCIENCE - Invasives/SPECIES/Whirling Disease/Monitoring/WD_sampling_results_fish_eDNA_used_for_making_maps_CMADSEN.xlsx"),
            to = paste0('data/WD_sampling_results_fish_eDNA_used_for_making_maps_CMADSEN.xlsx'),
            overwrite = T)
}

dat = read_excel('data/WD_sampling_results_fish_eDNA_used_for_making_maps_CMADSEN.xlsx', sheet = "Fish and eDNA")

dat = purrr::set_names(dat, snakecase::to_snake_case)

dat = dat |> filter(!is.na(lat) & !is.na(long))

# Split rows that have both eDNA and fish sampling into two rows each.
dat = dat |>
  tidyr::separate_longer_delim(cols = sampling_method, delim = " + ")

dat = sf::st_as_sf(dat, coords = c("long","lat"), crs = 4326)

# Typo corrections etc.
dat = dat |>
  dplyr::mutate(fish_sampling_results_q_pcr_mc_detected = ifelse(str_detect(fish_sampling_results_q_pcr_mc_detected,"Positive"),"Positive",fish_sampling_results_q_pcr_mc_detected)) |>
  dplyr::mutate(comments = ifelse(comments == '', NA, comments))

# --------------------------

col = sf::read_sf("W:/CMadsen/shared_data_sets/Columbia_River_Big_Watershed.shp") |>
  dplyr::summarise() |>
  sf::st_transform(4326)

sf::write_sf(col, "data/columbia_watershed.gpkg")

# ---------------------------
subw = sf::read_sf("W:/CMadsen/shared_data_sets/Columbia_River_Big_Watershed.shp") |>
  sf::st_transform(4326) |>
  dplyr::select(watershed_name = MAJOR_WA_1)


# lets get positive/negative for these results - not the tubifex
# the data has detected, weak detection, for edna+results_mc - change to pos/neg for now
dat_plotting<- dat |>
  plyr::mutate(e_dna_results_mc = ifelse(str_detect(e_dna_results_mc,"Weak Detection"),"Positive","Negative"))

dat_plotting$geometry[which(sf::st_coordinates(dat_plotting)[,1] == 15.0819)] <-
  sf::st_sfc(sf::st_point(c(-115.0819, 49.30667)), crs = 4326) # ask Chris about this one
# #leaflet map of the locations - we can just look at positive and negative for e_dna_results_mc or fish_sampling_results_q_pcr_mc_detected
# leaflet(dat_plotting) |> 
#   addTiles() |> 
#   addPolygons(data = col, fillColor = "transparent", color = "black", weight = 2) |> 
#   addCircleMarkers(
#     data = dat_plotting,
#     radius = 3,
#     color = ~ifelse(e_dna_results_mc == "Positive", "red", "blue"),
#     stroke = FALSE,
#     fillOpacity = 0.8,
#     popup = leafpop::popupTable(
#       dat_plotting,
#       zcol = c("sampling_method", "e_dna_results_mc", "fish_sampling_results_q_pcr_mc_detected")
#     )
#   ) |> 
#   addLegend(
#     position = "bottomright",
#     colors = c("red", "blue"),
#     labels = c("Positive", "Negative"),
#     title = "Sampling Results"
#   )


# Snag data directly from WD dashboard

# if(interactive()){
#   wd_res = sf::read_sf("../WhirlingDiseaseDashboard/app/www/sampling_results.gpkg")
#   col_ws = sf::read_sf("../WhirlingDiseaseDashboard/app/www/columbia_watershed.gpkg")
# } else {
#   wd_res = sf::read_sf("../../../WhirlingDiseaseDashboard/app/www/sampling_results.gpkg")
#   col_ws = sf::read_sf("../../../WhirlingDiseaseDashboard/app/www/columbia_watershed.gpkg")
# }


```

#### Assess watershed groups with parasite results

Either positive fish samples or suspect eDNA results.

```{r subw_with_positive_results}

subw = sf::read_sf("data/freshwater_atlas_watershed_groups_simplified.gpkg")
# for the points where there are positives, lets create a 2km buffer around them and 
# then filter the fish occurrences to only those that are within that buffer.
# buffer_distance <- 20000

# Transform to a projected CRS suitable for buffering
dat_plotting_crs <- sf::st_transform(dat_plotting, crs = 3005)

# Filter to points with either eDNA or fish sampling positives
positive_points <- dat_plotting_crs |>
  dplyr::filter(e_dna_results_mc == "Positive" | fish_sampling_results_q_pcr_mc_detected == "Positive")

# Join the positive points to watergroup groups
subw_pos = subw |> 
  sf::st_filter(sf::st_transform(positive_points,4326))

# Create the buffer around those points
# buffered_points <- sf::st_buffer(positive_points, dist = buffer_distance)

#ggplot()+geom_sf(data = buffered_points)


# buffered_points <- sf::st_transform(buffered_points, crs = 4326)
fish_occs_buffered <- sf::st_filter(fish_occs, subw_pos)
if('geometry' %in% names(fish_occs_buffered)){
  fish_occs_buffered = fish_occs_buffered |> 
    dplyr::rename(geom = geometry)
}
fish_occs_buffered <- fish_occs_buffered |> 
  dplyr::filter(!is.na(geom)) |> 
  dplyr::select(Species, geom, Location)

bbox <- st_bbox(fish_occs_buffered)

ggplot() +
  geom_sf(data = bc, color = "lightgrey") + 
  geom_sf(data = subw_pos, fill = NA, color = "darkgrey", size = 0.5) +
  geom_sf(data = fish_occs_buffered, aes(color = Species), size = 0.75) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(title = "Fish Occurrences within the Watershed Groups with Positive Sampling Results") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  coord_sf(
    xlim = c(bbox["xmin"] - 0.1, bbox["xmax"]+0.1),
    ylim = c(bbox["ymin"] -0.1, bbox["ymax"]+0.1)
  )

```

### Buffering by temperature restrictions

```{r buffer_filter_hydroclim}
# now that we have these points - we should filter by each of the temperatures!
# Extract the monthly temperatures for the buffered fish occurrences
extracted_values_buffered <- terra::extract(min_hydroclim_stack, fish_occs_buffered)
# Add fish_id for merging
fish_occs_buffered$fish_id <- 1:nrow(fish_occs_buffered)
# Merge the extracted temperature values with the spatial data
fish_temps_buffered_wide <- merge(
  as.data.frame(fish_occs_buffered),
  extracted_values_buffered,
  by.x = "fish_id",
  by.y = "ID",
  all.x = TRUE
)
# Rename the monthly temperature columns
names(fish_temps_buffered_wide)[grep("bc_Min.monthlyairtemperaturefor", names(fish_temps_buffered_wide))] <- month_names
# Pivot to long format
fish_temps_buffered_long <- pivot_longer(
  fish_temps_buffered_wide,
  cols = all_of(month_names),
  names_to = "month",
  values_to = "temperature"
)
# Filter by temperature range
fish_temps_buffered_long <- fish_temps_buffered_long %>%
  # filter(temperature >= 10, temperature <= 17)
  filter(temperature >= 10)

# Convert to sf using saved lon/lat
#fish_temps_buffered_long <- sf::st_as_sf(fish_temps_buffered_long, coords = c("lon", "lat"), crs = 4326)
# Set month factor levels for plotting
fish_temps_buffered_long$month <- factor(
  fish_temps_buffered_long$month,
  levels = c("January", "February", "March", "April", "May", "June",
             "July", "August", "September", "October", "November", "December")
)
# Plot
fish_temps_buffered_long <- sf::st_sf(fish_temps_buffered_long, sf_column_name = "geom")
bbox <- st_bbox(fish_temps_buffered_long)

ggplot() +
  geom_sf(data = bc, color = "lightgrey") +
  geom_sf(data = fish_temps_buffered_long, aes(color = Species), size = 0.75) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~ month) +
  theme_minimal() +
  labs(
    title = "Buffered Fish Occurrences with Temperatures Above 10째C Air Temperature",
    subtitle = "Faceted by Month"
  )+
   guides(color = guide_legend(override.aes = list(size = 2)))+
  coord_sf(
    xlim = c(bbox["xmin"] - 0.1, bbox["xmax"]+0.1),
    ylim = c(bbox["ymin"] -0.1, bbox["ymax"]+0.1)
  )

fish_temps_buffered_long |> 
  dplyr::select(Species, month, temperature, Location) |> 
  dplyr::mutate(temperature = round(temperature, 2)) |>
  st_drop_geometry() |> 
  DT::datatable(
    options = list(pageLength = 10, autoWidth = TRUE),
    rownames = FALSE,
    filter = "top",
    extensions = "Buttons",
    escape = FALSE,
    class = "display"
  ) 


```

#### Waterbodies within the subwatersheds with positive results

These are the named waterbodies within the subwatersheds where there have been positive fish results or suspected eDNA results.

```{r watershed waterbody overlap}

#get waterbodies within the subwatersheds with positive results

ws_rivers <- bcdata::bcdc_query_geodata('freshwater-atlas-rivers') |>
  filter(`WATERSHED_GROUP_CODE` %in% subw_pos$WATERSHED_GROUP_CODE) |> 
  collect() |> 
  st_transform(4326)

ws_lakes<-bcdata::bcdc_query_geodata('freshwater-atlas-lakes') |>
  filter(`WATERSHED_GROUP_CODE` %in% subw_pos$WATERSHED_GROUP_CODE) |> 
  collect()|> 
  st_transform(4326)

bbox<- st_bbox(ws_rivers)

ws_rivers <- ws_rivers |>  mutate(Type = "River") |> 
  filter(!is.na(GNIS_NAME_1))
ws_lakes <- ws_lakes |>  mutate(Type = "Lake") |> 
  filter(!is.na(GNIS_NAME_1))

ws_waterbodies<-rbind(ws_rivers, ws_lakes)

ggplot() +
  geom_sf(data = bc, color = "lightgrey") +
  geom_sf(data = ws_rivers, aes(color = Type), size = 0.75, show.legend = TRUE) +
  geom_sf(data = ws_lakes, aes(color = Type), fill = "purple", alpha = 0.5, show.legend = TRUE) +
  scale_color_manual(
    name = "Waterbody Type",
    values = c("River" = "blue", "Lake" = "purple")
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        fill = c("blue", "purple"),  # Fill the legend boxes with the same colors
        size = 4                     # Optional: make the legend symbols more visible
      )
    ),
    fill = "none"  # Hide the fill legend
  ) +
  theme_minimal() +
  labs(
    title = "Waterbodies within Subwatersheds with Positive Results"
  ) +
  coord_sf(
    xlim = c(bbox["xmin"] - 0.1, bbox["xmax"] + 0.1),
    ylim = c(bbox["ymin"] - 0.1, bbox["ymax"] + 0.1)
  )



ws_waterbodies |>
  sf::st_drop_geometry() |> 
  select(WATERSHED_GROUP_CODE, AREA_HA, GNIS_NAME_1, Type) |> 
  rename(
    Watershed_Group_Code = WATERSHED_GROUP_CODE,
    Area_Hectares = AREA_HA,
    Waterbody_Name = GNIS_NAME_1,
    Type = Type) |> 
  DT::datatable(
    options = list(pageLength = 10, autoWidth = TRUE),
    rownames = FALSE,
    filter = "top",
    extensions = "Buttons",
    escape = FALSE,
    class = "display"
  )


```




```{r}
# #### Species with high susceptibility to whirling disease
# 
# These species have been deemed to have the highest susceptibility are widely distributed throughout the province
# 
# ```{r high_priority_species}
# 
# high_priority_species <- c(
#   "Sockeye Salmon", "Rainbow Trout", "Cutthroat Trout"
# )
# 
# selected_sp <- fish_occs |> 
#   filter(Species %in% high_priority_species)
# 
# 
# ggplot() +
#   geom_sf(data = bc, color = "lightgrey") + 
#   geom_sf(data = selected_sp, aes (color = Species), size = 0.75) +
#   scale_color_brewer(palette = "Dark2") +
#   facet_wrap( ~ Species)+
#   theme_minimal()+
#   labs(title = "High Priority Fish Species Occurrences in BC",
#        subtitle = "Species: Sockeye Salmon, Rainbow (Steelhead) Trout, Cutthroat Trout")
# 
# 

#```
```
