---
title: "Prioritization of Water Bodies for Whirling Disease Sampling"
author: "Chris Madsen & John Phelan"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(readxl)
library(BAMMtools)
library(terra)
# devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)


base_dir = stringr::str_extract(getwd(),"C:\\/Users\\/[a-zA-Z]+")
onedrive_wd = paste0(str_extract(getwd(),"C:/Users/[A-Z]+/"),"OneDrive - Government of BC/data/")
lan_root = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/"

# This function finds the natural breaks (or, "Jenks") for a numeric column
bin_to_natural_breaks = function(dat,variable){
  # find natural breaks ("jenks")
  nbreaks = BAMMtools::getJenksBreaks(var = dat[[variable]], k = 4)
  dat |> 
    dplyr::rename(var_to_bin = !!rlang::sym(variable)) |> 
    dplyr::mutate(var_bin = as.numeric(cut(var_to_bin,nbreaks))) |> 
    dplyr::mutate(var_bin = tidyr::replace_na(var_bin, 1)) |> 
    dplyr::mutate(var_bin = as.character(var_bin)) |> 
    dplyr::mutate(var_bin = factor(var_bin, levels = c(1:3))) |> 
    dplyr::rename(!!rlang::sym(paste0(variable,"_bin")) := var_bin,
                  !!rlang::sym(variable) := var_to_bin)
}

## GGplot label convenience function
bin_info_labs = function(bins_title){
    labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)',
       subtitle = paste0("Inspection Bins: 1 ('low', ",
                         bins_title[1]," to ",bins_title[2],
                         "), 2 ('medium', ",
                         bins_title[2]," to ",bins_title[3],
                         "), 3 ('high', ",
                         bins_title[3]," to ",bins_title[4],
                         ")"))
}
```

### Watercraft Inspections Coming from Infected States/Provinces

```{r watercraft_insp_from_inf_areas}
insp_from_wd_inf_to_wb = sf::read_sf("W:/CMadsen/Projects/ZQMussels/data/Waterbodies_with_Inspection_Data_Summaries_all_years_WD_Infected_Areas.gpkg")
# Note: the above file can kind of only be modified / updated on Chris' computer, as it depends on the ZQMussels R project (specifically, it's Options.csv file) and the {imdp} R package... but the output gets saved to Chris' W drive, so that should be accessible for everyone!

# Tentative # 
# We could maybe use this version / cut of waterbodies in BC, add to it at each step, and that could be our list of waterbodies? #

insp_from_wd_inf_to_wb_b = bin_to_natural_breaks(insp_from_wd_inf_to_wb, 'TotalInspections')

# Strip out pacific ocean and 'dry storage' destinations for watercraft.
insp_from_wd_inf_to_wb_b = insp_from_wd_inf_to_wb_b |> 
  dplyr::filter(!GNIS_NA %in% c("Pacific Ocean","Dry Storage"))

bins_title = BAMMtools::getJenksBreaks(var = insp_from_wd_inf_to_wb$TotalInspections, k = 4)

# Plot showing binned numbers of watercraft headed to waterbodies in BC from infected states / provinces as well as infected waterbodies in BC.
ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = insp_from_wd_inf_to_wb_b, aes(fill = TotalInspections_bin,
                                             col = TotalInspections_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)',
       subtitle = paste0("Inspection Bins: 1 ('low', ",
                         bins_title[1]," to ",bins_title[2],
                         "), 2 ('medium', ",
                         bins_title[2]," to ",bins_title[3],
                         "), 3 ('high', ",
                         bins_title[3]," to ",bins_title[4],
                         ")"))

wb_list = insp_from_wd_inf_to_wb_b |> 
  dplyr::select(GNIS_NA, WATERSH, TotalInspections, TotalInspections_bin)
```

Currently, the number of distinct waterbodies with information is `r nrow(wb_list)`.

### Watercraft Inspections Coming from Infected Waterbodies in BC

```{r watercraft_insp_from_inf_wbs_in_bc}
insp_from_wd_bc_wb = sf::read_sf("W:/CMadsen/Projects/ZQMussels/data/Waterbodies_with_Inspection_Data_Summaries_all_years_BC_WD_Infected_Waterbodies.gpkg")
# Note: the above file can kind of only be modified / updated on Chris' computer, as it depends on the ZQMussels R project (specifically, it's Options.csv file) and the {imdp} R package... but the output gets saved to Chris' W drive, so that should be accessible for everyone!

insp_from_wd_bc_wb_b = bin_to_natural_breaks(insp_from_wd_bc_wb, 'TotalInspections')

# Strip out pacific ocean and 'dry storage' destinations for watercraft.
insp_from_wd_bc_wb_b = insp_from_wd_bc_wb_b |> 
  dplyr::filter(!GNIS_NA %in% c("Pacific Ocean","Dry Storage"))

bins_title_bc = BAMMtools::getJenksBreaks(var = insp_from_wd_bc_wb_b$TotalInspections, k = 4)

# Plot showing binned numbers of watercraft headed to waterbodies in BC from infected states / provinces as well as infected waterbodies in BC.
ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = insp_from_wd_bc_wb_b, aes(fill = TotalInspections_bin,
                                             col = TotalInspections_bin)) + 
  ggthemes::theme_map() + 
  scale_fill_brewer(palette = 'Spectral', direction = -1) +
  scale_color_brewer(palette = 'Spectral', direction = -1) + 
  labs(color = 'Inspections (bin)',
       fill = 'Inspections (bin)',
       subtitle = paste0("Inspection Bins: 1 ('low', ",
                         bins_title_bc[1]," to ",bins_title_bc[2],
                         "), 2 ('medium', ",
                         bins_title_bc[2]," to ",bins_title_bc[3],
                         "), 3 ('high', ",
                         bins_title_bc[3]," to ",bins_title_bc[4],
                         ")"))

# Identify any water bodies that aren't already in the list.
new_wb_for_list = insp_from_wd_bc_wb_b |> 
  dplyr::anti_join(
    wb_list |> 
      sf::st_drop_geometry() |> 
      dplyr::select(GNIS_NA, WATERSH)
  ) |> 
  dplyr::select(GNIS_NA, WATERSH,
                insp_from_wd_bc_wb = TotalInspections,
                insp_from_wd_bc_wb_b = TotalInspections_bin)

wb_already_in_list = insp_from_wd_bc_wb_b |> 
  dplyr::filter(paste0(WATERSH,GNIS_NA) %in% paste0(wb_list$WATERSH,wb_list$GNIS_NA)) |> 
  dplyr::select(GNIS_NA, WATERSH, insp_from_wd_bc_wb = TotalInspections,
                insp_from_wd_bc_wb_b = TotalInspections_bin) |> 
  sf::st_drop_geometry()

# Not new waterbodies per list, per se, but new columns.
wb_list = wb_list |> 
  dplyr::left_join(
    wb_already_in_list
  )

# Are there any new waterbodies to be added to the wb_list?
if(nrow(new_wb_for_list) > 0){
  wb_list = wb_list |> 
    dplyr::bind_rows(
      new_wb_for_list
    )
}
```

Currently, the number of distinct waterbodies with information is `r nrow(wb_list)`.

### Freshwater Fisheries Society Angler Survey (2022-2023)
```{r}
# We do have angler use as a raster on onedrive, but we also have the original polygon object on the W: drive.
# ang = terra::rast(paste0(onedrive_wd,"CNF/DFO_angling_survey_days_fished_raster.tif"))
ang = sf::read_sf("W:/CMadsen/shared_data_sets/freshwater_fisheries_society_angler_survey_2022_2023.gpkg")

# Drop waterbodies with no fishing
ang = ang |> 
  dplyr::filter(!is.na(days_fished))

# Rename some columns.
ang = ang |> 
  dplyr::rename(WATERSH = WATERSHED_GROUP_ID,
                GNIS_NA = Waterbody)

# Summarise at the waterbody name and watershed number level.
ang_s = ang |> 
  dplyr::group_by(GNIS_NA, WATERSH) |> 
  dplyr::summarise(days_fished = sum(days_fished,na.rm=T)) |> 
  dplyr::ungroup()

ang_s = bin_to_natural_breaks(ang_s, "days_fished")

bins_title_ang = BAMMtools::getJenksBreaks(ang_s$days_fished, k = 4)

ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = ang_s, aes(col = days_fished_bin, fill = days_fished_bin)) + 
  ggthemes::theme_map() + 
  bin_info_labs(bins_title_ang)

# Identify any water bodies that aren't already in the list.
new_wb_for_list = ang_s |> 
  dplyr::anti_join(
    wb_list |> 
      sf::st_drop_geometry() |> 
      dplyr::select(GNIS_NA, WATERSH)
  ) |> 
  dplyr::select(WATERSH, GNIS_NA, days_fished, days_fished_bin)

wb_already_in_list = ang_s |> 
  dplyr::filter(paste0(WATERSH,GNIS_NA) %in% paste0(wb_list$WATERSH,wb_list$GNIS_NA)) |> 
  dplyr::select(GNIS_NA, WATERSH, days_fished, days_fished_bin) |> 
  sf::st_drop_geometry()

# Not new waterbodies per list, per se, but new columns.
wb_list = wb_list |> 
  dplyr::left_join(
    wb_already_in_list
  )

# Are there any new waterbodies to be added to the wb_list?
if(nrow(new_wb_for_list) > 0){
  wb_list = wb_list |> 
    dplyr::bind_rows(
      new_wb_for_list
    )
}
```

Currently, the number of distinct waterbodies with information is `r nrow(wb_list)`.

### Hatchery Locations

```{r hatchery raster}
hatchery_rast<-terra::rast(paste0(onedrive_wd,"raster/aquaculture_presence_absence_raster.tif"))
#plot

hatchery_points <- terra::as.points(hatchery_rast, na.rm = TRUE)

hatchery_sf <- sf::st_as_sf(hatchery_points)

# Filter to only presence points (value = 1)
hatchery_presence_sf <- hatchery_sf[hatchery_sf$sum == 1, ] |> 
  dplyr::rename(hatchery_present = sum)

# Plot with a single color and simplified legend
ggplot() +
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = hatchery_presence_sf, color = "darkred", size = 1) +
  ggthemes::theme_map() +
  labs(color = NULL, subtitle = "Hatchery Presence in BC") +
  theme(legend.position = "none")

# Add in overlap of hatcheries to wb_list
hatcheries_by_wb = hatchery_presence_sf |> 
      sf::st_transform(3005) |> 
  sf::st_join(wb_list |> dplyr::select(GNIS_NA,WATERSH)) |> 
  sf::st_drop_geometry() |> 
  dplyr::filter(!is.na(GNIS_NA)) |> 
  dplyr::group_by(GNIS_NA, WATERSH) |> 
  dplyr::reframe(hatchery_present = sum(hatchery_present))

wb_list = wb_list |> 
  dplyr::left_join(
    hatcheries_by_wb
  )
```

### Watershed Groups with Positive Fish eDNA Detection
```{r}
ws = sf::read_sf("W:/CMadsen/shared_data_sets/subwatersheds_BC.shp")

pos_fish_edna = readxl::read_excel("J:/2 SCIENCE - Invasives/SPECIES/Whirling Disease/Monitoring/WD_sampling_results_fish_eDNA_used_for_making_maps_CMADSEN.xlsx") |> 
  purrr::set_names(snakecase::to_snake_case)

pos_fish_edna = pos_fish_edna |> 
  dplyr::filter(stringr::str_detect(fish_sampling_results_q_pcr_mc_detected, 'Positive'))

pos_fish_edna_sf = sf::st_as_sf(pos_fish_edna,coords = c('long','lat'), crs = 4326)

# Also pull in emerald lake, find its centroid.
emerald_lake = bcdata::bcdc_query_geodata('freshwater-atlas-lakes') |> 
  bcdata::filter(GNIS_NAME_1 == 'Emerald Lake', WATERBODY_POLY_ID == 705013945) |> 
  collect()

emerald_lake_centroid = emerald_lake |> 
  sf::st_centroid()

ws_w_edna_fish = ws |> 
  sf::st_filter(
    pos_fish_edna_sf |> 
      sf::st_transform(3005) |> 
      dplyr::bind_rows(emerald_lake_centroid)
  )

ws_bbox = sf::st_bbox(ws_w_edna_fish |> sf::st_buffer(dist = 50000))

ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) + 
  geom_sf(data = ws_w_edna_fish, aes(fill = "Positive Fish \neDNA Results")) +
  geom_sf(data = pos_fish_edna_sf, col = 'lightgreen') + 
  geom_sf(data = emerald_lake, fill = 'lightgreen', col = 'lightgreen') + 
  geom_sf_label(data = emerald_lake, aes(label = 'Emerald Lake'), nudge_x = 80000) +
  # geom_sf_label(data = pos_fish_edna_sf, aes(label = sample_site_name), nudge_x = 80000, nudge_y = -100000) +
  ggsflabel::geom_sf_label_repel(data = pos_fish_edna_sf,
                                 aes(label = sample_site_name),
                        force = 10, seed = 10) +
  labs(fill = 'eDNA') + 
  coord_sf(xlim = ws_bbox[c(1,3)],
           ylim = ws_bbox[c(2,4)]) + 
  ggthemes::theme_map() + 
  ggspatial::annotation_scale()

wb_list = wb_list |> 
  dplyr::mutate(fish_edna_positive_in_watershed = WATERSH %in% ws_w_edna_fish$WATERSHED_)
```

			vi. Overlay eDNA results for tubifex
			vii. If tubifex density is accurate, overlay tubifex density map binned into presence/absence (doesn't contribute to score)
			viii. Overlay susceptible species occurrences (present/absent for each waterbody) and facet map by species at this stage; maybe later we'll remove WBs with 0. We could apply this and show the effect of doing this.
			ix. Add in expert opinion from regions, First Nations, etc., as available
			x. In future, we'll bring in SARA-listed species / COSEWIC

```{r add_together_variables}
wb_list = wb_list |> 
  dplyr::mutate(priority = TotalInspections_bin + insp_from_wd_bc_wb_b + days_fished_bin)
```

```{r show_results_data_table}
wb_list |> 
  sf::st_drop_geometry() |> 
  DT::datatable()
```
